---
title: "R Notebook"
output: html_notebook
---

#3.2 The very basics

#3.2.1 Objects
When solving quadratic equations, the solutions change depending on the values of a, b and c. Programming language can define variables and write expressions with these variables.

```{r}
a <- 1
b <- 1
c <- -1
```

We can also use = , but <- is recommended.

To see the value stored in variable,
```{r}
a
```
A more explicit way
```{r}
print(a)
```

We use the term 'object' to describe stuff that is stored in R. Objects can also be functions.

#3.2.2 The workspace

As we define objects in the console, we are actually changing the workspace.
```{r}
ls()
```
the Environment tab shows the values.

```{r}
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```


#Functions

The data analysis process can usually be described as a series of functions applied to the data. R includes predefined functions. They are available for immediate use. Unlike ls, most functions require one or more arguments. 

```{r}
log(8)
log(a)
```

```{r}
help("log")
?log
```
The help page will show you what arguments the function is expecting. You can determine which arguments are optional by noting in the help document that a default value is assigned with =. 

If you want a quick look at the arguments without opening the help system,
```{r}
args(log)
```

You can change the default values by
```{r}
log(8, base = 2)
log(x=8, base=2)
```

by not using the names, it assumes the arguments are x, base
```{r}
log(8,2)
```

If using the argumenst' names, then we can include them in whatever order we want
```{r}
log(base =2, x=8)
```
To specify arguments, we must use =, and cannot use <-.

```{r}
2^3
```

```{r}
help("+")
help("<")
```

#Other prebuilt objects

You can see available datasets by
```{r}
data()
```
```{r}
co2
pi
Inf+1
```

#Variable names

Some basic rules in R: variable names have to start with a letter, can't contain spaces and should not be predefined variables.
Convention: use meaningful words, use only lower case, and use underscores(_) as a substitute for spaces.
For example, 
```{r}
solution_1 <- (-b + sqrt(b^2 - 4*a*c)) / (2*a)
solution_2 <- (-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

# Saving your workspace

Values remain in the workspace until you end or erase them with the function 'rm'. 

Assign the workspace a specific name. To load, use the function 'load'. When saving a workspace, use the suffix 'rda' or 'RData'.

# Motivating scripts

To solve another equation, we can copy and paste the code and redefine the variables and recompute the solution
```{r}
a <- 3
b <- 2
c <- -1
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

# Commenting your code

If a line of R code starts with the symbol #, it is not evaluated. We can use this to write reminders of why we wrote particular code.
For example,
```{r}
##Code to compute solution to quadratic equation of the form ~~
##define the variables
a <- 3 
b <- 2
c <- -1

## now compute the solution
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```









